Assessment of the KLGO Coastal Waterbird Survey Monitoring Program

### 2019-02-21 analysis/paper/paper.Rmd
Example of next code chunks to work on in paper.Rmd, and modified survey_timing to clarify what next steps are needed.

[@Hahr2004], pg 13: Areas of coastline that were visible from public access points
between Skagway and the Taiya River delta were divided into eight census units
and surveyed once per week in late April and May and bi-weekly during late
summer and fall (August – mid September) (Figure 5). Surveys were conducted
as weather and scheduling permitted and dates/time were not selected at
random. Standardized observation points were established for each census unit
to ensure the most complete coverage of the study area, and to minimize
observer bias.
Observers counted and identified all birds visible within each unit and
documented birds opportunistically observed while moving between units.
Counts were made by scanning the census units with Bushnell Legend 10 x 42
binoculars and a Bausch & Lomb 20-60 x 70 spotting scope until the observer
was able to quantify all visible birds in the unit. Surveys were only conducted
under optimal weather conditions; however, occasionally conditions deteriorated
during the course of a survey making species identification difficult (especially in
the case of low clouds, fog and wind). Data were recorded on standardized data
forms developed after Collins et al. (2001) (Appendix B). ArcView shapefiles of
the observation points and census units were created for use in the park’s GIS.
All data, original data forms, and field notes were archived at KLGO.

@@

>>>>

[@Hahr2004], pg 13: Areas of coastline that were visible from public access points
between Skagway and the Taiya River delta were divided into eight census units
and surveyed once per week in late April and May and bi-weekly during late
summer and fall (August – mid September) (Figure 5). Surveys were conducted
as weather and scheduling permitted and dates/time were not selected at
random. Standardized observation points were established for each census unit
to ensure the most complete coverage of the study area, and to minimize
observer bias.

Observers counted and identified all birds visible within each unit and
documented birds opportunistically observed while moving between units.
Counts were made by scanning the census units with Bushnell Legend 10 x 42
binoculars and a Bausch & Lomb 20-60 x 70 spotting scope until the observer
was able to quantify all visible birds in the unit. Surveys were only conducted
under optimal weather conditions; however, occasionally conditions deteriorated
during the course of a survey making species identification difficult (especially in
the case of low clouds, fog and wind). Data were recorded on standardized data
forms developed after Collins et al. (2001) (Appendix B). ArcView shapefiles of
the observation points and census units were created for use in the park’s GIS.
All data, original data forms, and field notes were archived at KLGO.

@@

Narrative on how well the current surveys are capturing the Spring arrival and Fall departure of each of the targeted species (\@ref(fig:Fig-SpeciesTiming)). Implications for informing Audubon Climate Change Watch or national phenology network, yadda yadd.

```{r SpeciesTiming, eval=TRUE, cache = FALSE, include = FALSE}
# Madeleine: write a script to extract, for just the target species listed at line 250, 
# just the dates each year that each of the target species was observed (somewhere in the survey areas - don't care which, at the moment)
# then add a layer to the ggplot 'Plot1' (horrid name, by the way) to visually distinguish those dates for that species.
# will have to play around to see if it is feasible to do this by faceting on the species or just make a separate 
# timing graph for each species - your call. It will be unreadable if try and put all of the target species on 
# a single plot

# tweak this call to source that new script, then un-comment, etc.
# source("../scripts/species_timing_EDA.R", echo=FALSE)

```{r Fig-SpeciesTiming, echo=FALSE, eval=TRUE, cache=TRUE, fig.show= 'asis', fig.cap="For each of the target species, the dates on which they were observed in any of the sample areas."}

#edit to use the name of the new plot
print(Plot1)
```
